Virtual Memory Assisted Transactional Memory for Dynamic Languages

### 2014-06-04 talk/dls2014/paper/paper.tex
simply move memory requirements explanation to its own section in low-level impl

@@

\cfbolz{part of this section is wrong here. maybe in the low-level section? here you should only measure}

Second, there are areas in memory private to each segment (see
Section~\ref{sub:Setup}). The Nursery for example is 4~MiB in
size. Additionally, the area used for the read markers is lazily
mapped by the kernel and only needed for objects in the old object
space. Thus, it can be at most the total size of memory used by
old objects divided by 16.

Third, since we do copy on write, if all objects were to be written to
at the same time, all pages would need to be privatised for all
objects in the old object space. In that case we would need the total
amount of memory required by old objects multiplied by $N+1$
(incl. the sharing segment). During major collections we re-share the
pages if possible.

\begin{figure}[h]
  \centering
  \includegraphics[width=1\columnwidth]{plots/richards_mem.pdf}
  \caption{Actual memory managed by the GC and the page privatisation
    over time in Richards benchmark\label{fig:richards_mem}}
\end{figure}

In Figure~\ref{fig:richards_mem} we look at the memory usage of one of
our benchmarks called Richards.\footnote{Richards is a OS kernel simulation
benchmark.} The \emph{GC managed memory} counts all memory used in the
old object space including the memory required for private pages. Note,
it does not measure the memory required for the segment-private areas
like the nursery and read markers. The
sharp drops in memory usage come from major collections that free old
objects and re-share pages. The average memory usage stays around
29~MiB and we see that in this benchmark we have around 1 major
collection every second.

The \emph{page privatisation}, which represents the ratio between
private pages and shared pages, gives us an idea about how many
private copies exist for each shared page. If the ratio is $>1.0$, it
means that on average we have more than one private copy for each
shared page. In the worst case described above, this number would
reach $N$ and means we need $N\times$ the memory for the private
pages alone. In this benchmark, we see the same spikes as the memory
usage. These come directly from re-sharing the pages. The maximum page
privatisation is around $0.9$ between major collections, thus the
private pages are responsible for a $90\%$ increase in the required
memory. Since the spikes in the GC managed memory line actually
show increases by $~100\%$, it means that the rest comes from actual
garbage objects that were collected.

>>>>

\subsection{Memory Requirements\label{sub:memoryreq}}

Second, there are areas in memory private to each segment (see
Section~\ref{sub:Setup}). The Nursery for example is 4~MiB in
size\footnote{Configurable value taken from experience with our
non-STM reference interpreter}. Additionally, the area used for the
read markers is lazily mapped by the kernel and only needed for
objects in the old object space. Thus, it can be at most the total
size of memory used by old objects divided by 16, adding up to
$6.3\%$ to the total memory requirements.

Third, since we do copy on write, if all objects were to be written to
at the same time, all pages would need to be privatised for all
objects in the old object space. In that case we would need the total
amount of memory required by old objects multiplied by $N+1$
(incl. the sharing segment). This is certainly the worst case and
so far not observed in practice. During major collections we re-share
privatised pages that are currently not modified by any transaction.

@@

\remi{reference \ref{sub:memoryreq}}

% \begin{figure}[h]
%   \centering
%   \includegraphics[width=1\columnwidth]{plots/richards_mem.pdf}
%   \caption{Actual memory managed by the GC and the page privatisation
%     over time in Richards benchmark\label{fig:richards_mem}}
% \end{figure}

% In Figure~\ref{fig:richards_mem} we look at the memory usage of one of
% our benchmarks called Richards.\footnote{Richards is a OS kernel simulation
% benchmark.} The \emph{GC managed memory} counts all memory used in the
% old object space including the memory required for private pages. Note,
% it does not measure the memory required for the segment-private areas
% like the nursery and read markers. The
% sharp drops in memory usage come from major collections that free old
% objects and re-share pages. The average memory usage stays around
% 29~MiB and we see that in this benchmark we have around 1 major
% collection every second.

% The \emph{page privatisation}, which represents the ratio between
% private pages and shared pages, gives us an idea about how many
% private copies exist for each shared page. If the ratio is $>1.0$, it
% means that on average we have more than one private copy for each
% shared page. In the worst case described above, this number would
% reach $N$ and means we need $N\times$ the memory for the private
% pages alone. In this benchmark, we see the same spikes as the memory
% usage. These come directly from re-sharing the pages. The maximum page
% privatisation is around $0.9$ between major collections, thus the
% private pages are responsible for a $90\%$ increase in the required
% memory. Since the spikes in the GC managed memory line actually
% show increases by $~100\%$, it means that the rest comes from actual
% garbage objects that were collected.

\remi{do this and RSS numbers for pypy-stm-nojit and normal pypy-nojit:}
